

# Example KDE boundary issues

Gene Leynes 2016-09-20

## Background, assumptions, and sources

The following map shows car thefts in Chicago for the first six months of 2016.  This data set was chosen for its size; it's relatively small, but also hetereogenous and provides good coverage throughout the city. 

**Data source**: https://data.cityofchicago.org/Public-Safety/Crime-Car-Thefts-first-6-months-of-2016/22s8-eq8h

**Direct download** https://data.cityofchicago.org/api/views/22s8-eq8h/rows.csv?accessType=DOWNLOAD

The density estimates  were developed using R's `KernSmooth` package written by Matt Wand and Brian Ripley, based on the book "Kernel Smoothing" by Wand, M.P. and Jones, M.C. (1995).  The estimate used a gridsize of 100 by 100, and bandwiths of .0045 for Longitude and .0068 for Latitude.

## Narrative

The map below shows both points and density estimates for auto thefts around the city.  **You can easily zoom in and scroll around to see more detail; use your mouse scroll wheel and click and drag to explore the map**.

Although you may make an argument to experiment with a lower bandwidth, I think the levels look reasonable throughout much of the city.



```{r}
library(KernSmooth)
```


```{r, message=FALSE, echo = FALSE, results = "hide"}

## IF RUNNING THIS FOR THE FIRST TIME PLEASE INSTALL THESE PACKAGES WHICH 
## WILL INSTALL OTHER PACKAGES WITHIN R AUTOMATICALLY

## THIS SHOULD EXECUTE AUTOMATICALLY, BUT YOU MAY NEED TO INSTALL MANUALLY
##
## You will need devtools (which is used to install geneorama from github)
## Then you'll need geneorama
##
## You also need Rcpp, which is a pretty common package, but for some reason
## it works better to install separately first.
##
## After this, geneorama's function "loadinstall_libraries" will load or install
## any missing libraries.
##


if(!"devtools" %in% rownames(installed.packages())){
    install.packages("devtools",
                     dependencies = TRUE,
                     repos = "http://cran.rstudio.com/")
}

if(!"Rcpp" %in% rownames(installed.packages())){
    install.packages("Rcpp",
                     dependencies = TRUE,
                     repos = "http://cran.rstudio.com/")
}


if(!"geneorama" %in% rownames(installed.packages()) |
   installed.packages()["geneorama","Version"] < "1.5.0"){
    devtools::install_github('geneorama/geneorama')
}

```


```{r, message=FALSE, echo = FALSE, results = "hide"}
library(geneorama)
loadinstall_libraries("knitr")
options(datatable.prettyprint.char=80L)
set_project_dir("wnv_map_demo")

## INITIALIZE
loadinstall_libraries(c("leaflet", "data.table", "sp", "rgdal", "KernSmooth", "RCurl"))
# library("maptools")

inurl <- "https://data.cityofchicago.org/api/views/22s8-eq8h/rows.csv?accessType=DOWNLOAD"
infile <- "booma_example/mvthefts.Rds"

## LOAD DATA
## Also, clean up variable names, and convert dates
if(!file.exists(infile)){
    cat("Downloading file\n")
    # download.file(url = inurl, destfile = infile)
    dat <- fread(RCurl::httpGET(inurl)[1])
    setnames(dat, tolower(colnames(dat)))
    setnames(dat, gsub(" ", "_", colnames(dat)))
    dat <- dat[!is.na(longitude)]
    dat[ , date := as.IDate(date, "%m/%d/%Y")]
    saveRDS(dat, infile)
}
dat <- readRDS(infile)

## MAKE CONTOUR LINES
## Note, bandwidth choice is based on MASS::bandwidth.nrd()
kde <- bkde2D(dat[ , list(longitude, latitude)],
              bandwidth=c(.0045, .0068), gridsize = c(100,100))
CL <- contourLines(kde$x1 , kde$x2 , kde$fhat)

## EXTRACT CONTOUR LINE LEVELS
LEVS <- as.factor(sapply(CL, `[[`, "level"))
NLEV <- length(levels(LEVS))

## CONVERT CONTOUR LINES TO POLYGONS
pgons <- lapply(1:length(CL), function(i)
    Polygons(list(Polygon(cbind(CL[[i]]$x, CL[[i]]$y))), ID=i))
spgons = SpatialPolygons(pgons)

## MAPBOX INFO
MAPBOX_STYLE_TEMPLATE <- paste0("https://api.mapbox.com/styles/v1/coc375492/",
                                "cirqd7mgf001ygcnombg4jtb4/tiles/256/{z}/{x}/{y}",
                                "?access_token=pk.eyJ1IjoiY29jMzc1NDkyIiwiYSI6ImN",
                                "pcnBldzVqMTBmc3J0N25rZTIxZ3ludDIifQ.DgJIcLDjC1h9MtT8CaJ-pQ")
mb_attribution <- paste("© <a href='https://www.mapbox.com/about/maps/'>Mapbox</a> ",
                        "© <a href='http://www.openstreetmap.org/about/'>OpenStreetMap</a>")

## CREATE LABEL
dat[ , LABEL := paste0(date, " | ", location_description, " | arrest:", arrest, 
                       " lon:", longitude, " lat:", latitude)]
## CREATE COLOR PALETTE
pal <- leaflet::colorFactor("Greens", levels = -NLEV:NLEV)
```


```{r, echo = FALSE}
## Little trick to make the legend look better:
i <- length(levels(LEVS))
legend_colors <- pal(round(seq(-i+3, i-3, length.out=i)))

leaflet(spgons) %>% 
    addTiles(urlTemplate = MAPBOX_STYLE_TEMPLATE, attribution = mb_attribution) 	%>%
    addPolygons(color = pal(as.numeric(LEVS)), weight=1, fillOpacity=.25) %>%
    addCircles(lng = ~longitude, lat = ~latitude, weight = 3, popup = ~LABEL,
               data = dat, radius = .5, opacity = .4, 
               col = ifelse(dat$arrest=="true", "yellow", "red")) %>% 
    addLegend(opacity = 1,
              # values = LEVS, pal = pal,
              labels = levels(LEVS),
              colors = legend_colors,
              position = "bottomleft", 
              title = "Crime Intensity") %>% 
    addLegend(colors = c("yellow", "red"),
              labels = c("true", "false"), 
              position = "bottomright", 
              title = "Arrest")
```


## Columbous Park Example 


The map below is the exact same map as the one above, but zoomed into an area of interest.  

This zoomed in map shows some problems with the KDE.  The KDE looks reasonable in areas that are completely internal. However, when there's an area at the periphery that area seems to be artificially low.

This area at Roosevelt and Austin Blvd is a relative hot spot, but it's level is only a 10. If you compare that with the point at Central and Madison, you'll see that a seemingly similar level of crime is reported as a 25. 


```{r, echo = FALSE}
leaflet(spgons) %>% 
    addTiles(urlTemplate = MAPBOX_STYLE_TEMPLATE, attribution = mb_attribution) 	%>%
    addPolygons(color = pal(as.numeric(LEVS)), weight=1, fillOpacity=.25) %>%
    addCircles(lng = ~longitude, lat = ~latitude, weight = 3, popup = ~LABEL,
               data = dat, radius = .5, opacity = .4, 
               col = ifelse(dat$arrest=="true", "yellow", "red")) %>% 
    addLegend(opacity = 1, 
              # pal = pal, values = LEVS, 
              labels = levels(LEVS),
              colors = legend_colors,
              position = "bottomleft", 
              title = "Crime Intensity") %>% 
    addLegend(colors = c("yellow", "red"),
              labels = c("true", "false"), 
              position = "bottomright", 
              title = "Arrest") %>%
    fitBounds(-87.81, 41.85, -87.734, 41.88)
```


